---
title: 'Reproducible Research: Peer Assessment 1'
author: "Emily Kubicek"
date: "12/9/2017"
output: md_document
---

*"This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day."*

Below are varying analyses conducted on the data starting with loading the data into R.

## Loading and preprocessing the data
#### 1. Load the data
#### 2. Process/transform the data
```{r}
library(knitr)
setwd("~/Desktop/Coursera/ReproducibleResearch")
data<-read.csv("activity.csv")
data2<- na.omit(data)
library(dplyr)
data3<-tbl_df(data2)
```

## What is mean total number of steps taken per day?
#### 1. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
teststeps <- tapply(data3$steps, data3$date, sum)
hist(teststeps, col = "lightblue", xlab = "Total Steps per Day", ylab = "Frequency", 
     main = "Histogram of Total Steps taken per day")

```

#### 2. Calculate and report the mean and median total of steps taken per day
```{r}
date_median<-aggregate(data3["steps"], by=data3["date"], median)
date_mean<-aggregate(data3["steps"], by=data3["date"], mean)
dmm<-inner_join(date_median, date_mean, by = "date")
names(dmm)[2:3] <- c("median","mean")
print(dmm)
```

## What is the average daily activity pattern?
#### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
mean_steps <- tapply(data3$steps, data3$interval, mean, na.rm = TRUE)
plot(row.names(mean_steps), mean_steps, type = "l", xlab = "Time Intervals (5-minute)", 
     ylab = "Mean of steps (all days)", 
     main = "Average Steps Taken at 5 minute Intervals", 
     col = "purple", lwd=2)
```

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
whichmax <- which.max(mean_steps)
maxsteps <- names(whichmax)
print("The interval with the highest number of steps is 835")
```

##Input missing values
#### 1. Calculate and report the total number of missing values in the data set.
```{r}
sum(is.na(data))
```

#### 2. Devise a strategy for filling in all of the missing values in the dataset.

#### 3. Create a new dataset that is equal to the original data set but with the missing values filled in.
```{r}
mean_steps <- tapply(data3$steps, data3$interval, mean, na.rm = TRUE)
head(data)
data[is.na(data)] <- mean_steps
newdata<- data
data<-read.csv("activity.csv")
head(newdata)
```

#### 4a. Make a histogram of the total number of steps taken each day
```{r}
nonas_teststeps <- tapply(newdata$steps, newdata$date, sum)
hist(nonas_teststeps, col = "darkblue", xlab = "Total Steps per Day", ylab = "Frequency", 
     main = "Histogram of Total Steps taken per day")
```

#### 4b. Calculate and report the mean/median total number of steps taken per day
```{r}
nonas_date_median<-aggregate(newdata["steps"], by=newdata["date"], median)
nonas_date_mean<-aggregate(newdata["steps"], by=newdata["date"], mean)
nonas_dmm<-inner_join(date_median, date_mean, by = "date")
names(nonas_dmm)[2:3] <- c("median","mean")
print(nonas_dmm)
```

#### 4c. Do these values differ from the estimates from the
#### first part of the assignment? What is the impact of
#### imputing missing data on the estimates of the total
#### daily number of steps?
##### As seen in the median and mean calculations for each individual day, these values do not change when missing values are replaced with the mean of the recorded steps for that interval.

## Are there differences in activity patterns between weekdays and weekends?
#### 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
library(dplyr)
library(lubridate)

testdata2 <- data3
# testdata2<-mutate(data3, day = weekdays(data3$date))

testdata2$date<-as.Date(data3$date)
testdata2$Weekend <- weekdays(testdata2$date)
testdata2$Weekend <- grepl("S.+",testdata2$Weekend)


testdata2$date <- as.Date(testdata2$date)
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
testdata2$day <- factor((weekdays(testdata2$date) %in% weekdays1), 
                   levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
testdata2
```

#### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
wdaydata<-subset(testdata2, testdata2$day=="weekday")
wdaymean<-tapply(wdaydata$steps, wdaydata$interval, mean, na.rm = TRUE)

wenddata<-subset(testdata2, testdata2$day=="weekend")
wendmean<-tapply(wenddata$steps, wenddata$interval, mean, na.rm = TRUE)


par(mfrow=c(2,1))
plot(names(wdaymean),wdaymean, type = "l",
     xlab = "Time Intervals (5-minute)",
     ylab = "Mean of steps",
     main = "Weekdays",
     col = "red", lwd=2)

plot(names(wendmean),wendmean, type = "l",
     xlab = "Time Intervals (5-minute)",
     ylab = "Mean of steps",
     main = "Weekends",
     col = "green", lwd=2)

```








